A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-18 survey.tex
Removed the paragraph about ``human complexity''

Another way  to measure the complexity  of the game is  by looking at
the branching factor, $b$, and the depth of the game, $d$, as proposed in~\cite{Gaby}, with a total game complexity of $b^d$. In Chess, $b \approx 35$ and $d \approx 80$. In more complex games, like Go, $b \approx 30 - 300$, and $d \approx 150 - 200$. In StarCraft, humans can issue only one action at a time (this is because of the GUI used by the game, computer players can issue several actions, to different units, at any given time), thus, assuming a player controls between 50 to 200 units, and that each unit can execute between 4 to 20 actions (moving in each of the 4 cardinal directions, attacking units in range, use special abilities, etc.), we estimate the branching factor for human players to be around $b \approx 200 - 4000$. Good players can execute about 300 actions per minute, and the typical length of a game is about 25 minutes. Thus, $d \approx 7500$.

>>>>

Another way  to measure the complexity  of the game is  by looking at
the branching factor, $b$, and the depth of the game, $d$, as proposed in~\cite{Gaby}, with a total game complexity of $b^d$. In Chess, $b \approx 35$ and $d \approx 80$. In more complex games, like Go, $b \approx 30 - 300$, and $d \approx 150 - 200$. In order to determine the branching factor in StarCraft when an AI plays it, we must have in mind, that the AI can issue actions simultaneously to as many units in the game as desired. Thus, considering that, in a typical game, a player controls between 50 to 200 units, the branching factor would be between $u^{50}$ and $u^{200}$, where $u$ is the number of actions each unit can execute. Assuming a conservative value for $u$ of about $30$ (moving to nearby positions, attacking units in range, using special abilities, etc.), we obtain a branching factor of $b$ between  $30^{50}$ and $30^{200}$. Now, to compute $d$, we simply consider the fact that typical games last for about 25 minutes, which results in $d \approx 36000$ (25 minutes $\times$ 60 seconds $\times$ 24 iterations per second).

%In StarCraft, humans can issue only one action at a time (this is because of the GUI used by the game, computer players can issue several actions, to different units, at any given time), thus, assuming a player controls between 50 to 200 units, and that each unit can execute between 4 to 20 actions (moving in each of the 4 cardinal directions, attacking units in range, use special abilities, etc.), we estimate the branching factor for human players to be around $b \approx 200 - 4000$. Good players can execute about 300 actions per minute, and the typical length of a game is about 25 minutes. Thus, $d \approx 7500$.
